---
title: "Datasaurus Dozen"
format: html
---

```{r}
#| echo: false
#| warning: false

# Pakete aktivieren ----
library(rio)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(scales)
library(gt)

# Get data from OSF
source("scripts/setup.R")

# Read data
datasaurus <- import("data/datasaurus.csv")
```

## Where the data comes from

> The Datasaurus dozen comprises thirteen data sets that have nearly identical simple descriptive statistics to two decimal places, yet have very different distributions and appear very different when graphed.[1] It was inspired by the smaller Anscombe's quartet that was created in 1973.
>
> -- [wikipedia.org](https://en.wikipedia.org/wiki/Datasaurus_dozen)

## What the data looks like

```{r}
datasaurus |> glimpse()
```

There are `{r} datasaurus$dataset |> unique() |> length()` data sets, with two variables each. See @tbl-summary for summary statistics.

> Similar to the Anscombe's quartet, the Datasaurus dozen was designed to further illustrate the importance of looking at a set of data graphically before starting to analyze according to a particular type of relationship, and the inadequacy of basic statistic properties for describing realistic data sets.
> 
> -- [wikipedia.org](https://en.wikipedia.org/wiki/Datasaurus_dozen)

```{r}
#| echo: false
#| warning: false
#| label: tbl-summary
#| tbl-cap: Summary statistics

datasaurus |> 
    group_by(dataset) |> 
    summarise(
        M_y = mean(y),
        SD_y = sd(y),
        M_x = mean(x),
        SD_x = sd(x),
        cor = cor(x, y)
    ) |> 
    gt() |> 
    fmt_number(M_y:SD_x) |> 
    fmt_number(cor, decimals = 3) |> 
    cols_label(
        M_y = "Mean y",
        M_x = "Mean x ",
        SD_y = "Std dev y",
        SD_x = "Std dev x",
        cor = "Correlation"
    )
```

## Plotting the data 

Looks like something is going on here …

```{r}
#| echo: false
#| warning: false
#| label: fig-plot-1
#| fig-cap: "All the data, no filters"
#| fig-alt: ""
#| fig-width: 8
#| fig-height: 6

datasaurus |> 
    ggplot(aes(x = x, y = y)) +
    geom_point() +
    labs(
        title = "The Datasauruns Dozen",
        subtitle = sprintf("All datasets have a mean of x: %.2f, mean of y: %.2f, cor(x, y): %.2f", mean(datasaurus$x), mean(datasaurus$y), cor(datasaurus$x, datasaurus$y)),
        x = "Dataset",
        y = "Mean"
    ) +
    theme(legend.position = "none")
```

Adding color helps, but it's still hard to read.

```{r}
#| echo: false
#| warning: false
#| label: fig-plot-2
#| fig-cap: "All the data, no filters"
#| fig-alt: ""
#| fig-width: 8
#| fig-height: 6

datasaurus |> 
    ggplot(aes(x = x, y = y, color = dataset)) +
    geom_point() +
    labs(
        title = "The Datasauruns Dozen",
        subtitle = sprintf("All datasets have a mean of x: %.2f, mean of y: %.2f, cor(x, y): %.2f", mean(datasaurus$x), mean(datasaurus$y), cor(datasaurus$x, datasaurus$y)),
        x = "Dataset",
        y = "Mean"
    )
```

Facets to the rescue! 

```{r}
#| echo: false
#| warning: false
#| label: fig-plot-3
#| fig-cap: "All the data, no filters"
#| fig-alt: ""
#| fig-width: 8
#| fig-height: 6

datasaurus |> 
    ggplot(aes(x = x, y = y, color = dataset)) +
    geom_point() +
    facet_wrap(~dataset) +
    labs(
        title = "The Datasauruns Dozen",
        subtitle = sprintf("All datasets have a mean of x: %.2f, mean of y: %.2f, cor(x, y): %.2f", mean(datasaurus$x), mean(datasaurus$y), cor(datasaurus$x, datasaurus$y)),
        x = "Dataset",
        y = "Mean"
    ) +
    theme(legend.position = "none")
```